---
title: "Topic 9 - Additional Slides"
subtitle: "These slides have also been included at teh end of the original topic 9 slides."
format: 
  revealjs:
    theme: OSU.scss
    chalkboard: false
    html-math-method: mathjax
    multiplex: true
editor: visual
---

```{r}
library(tidyverse)
```

## Download or print notes to PDF {.smaller}

If you'd like to export this presentation to a PDF, do the following

::: nonincremental
1.  Toggle into Print View using the E key.
2.  Open the in-browser print dialog (CTRL/CMD+P)
3.  Change the **Destination** to **Save as PDF**.
4.  Change the **Layout** to **Landscape**.
5.  Change the **Margins** to **None**.
6.  Enable the **Background graphics** option.
7.  Click **Save**.
:::

This feature has been confirmed to work in Google Chrome and Firefox.



```{r}
library(datasets)
 
state <- state.x77 |>
    as_tibble(rownames = "States")

set.seed(7493410)

keep <- sample(1:50, 30, replace = FALSE)

state_30 <- state[keep,]

```

```{r}
state_30 <- state_30 |>
  rename("LifeExp"=`Life Exp`)
```


```{r}

mod <- lm(LifeExp ~ Illiteracy, data = state_30)
```

## Prediction & Confidence Intervals {.smaller}

[The model predicted that for a state with an illiteracy rate $1.4\%$, the life expectancy was 70.577 years in 1970.]{style="color:blue;"}

[$$\hat{y} = 72.181 - 1.146(1.4) = 70.577$$]{style="color:blue;"}

::: fragment
#### Prediction Intervals for a Single Response Value

We know that this prediction in based on a sample and that a different sample of observations would have likely yielded a slightly different prediction.

[We can address the uncertainty in the prediction for this single response value using a **prediction interval**.]{.fragment}
:::

::: fragment
```{r}
#| echo: true
predict(mod, newdata = data.frame(Illiteracy = 1.4), interval = "prediction", level = 0.95)
```
:::

 

[For a single state with an illiteracy rate of 1.4% in 1970, we are 95% confident that the life expectancy was between 68.064 and 73.090 years, with a point estimate of 70.577 years.]{.fragment}

## Prediction & Confidence Intervals {.smaller}

The Central Limit Theorem tells us that mean values are more predictable than individual measurements.

::: fragment
Consider predicted the mean life expectancy for **all** states with an illiteracy rate of 1.4%.
:::

::: fragment
The estimate is the same as what we saw previously: [$\bar{y} = 72.181 - 1.146(1.4) = 70.577$]{style="color:blue;"}.
:::

:::: fragment
#### Confidence Intervals for a Mean Response Value

In addition to the point estimate for the mean life expectancy for **all** states with an illiteracy rate of 1.4%, we can provide a **confidence interval for the mean response value**.

::: fragment
```{r}
#| echo: true
predict(mod, newdata = data.frame(Illiteracy = 1.4), interval = "confidence", level = 0.95)
```

[We are 95% confident that mean life expectancy in 1970 for all states with an illiteracy rate of 1.4% was between 70.103 and 71.051 years, with a point estimate for the mean response of 70.577 years.]{.fragment}
:::
::::

## Prediction & Confidence Intervals {.smaller}

#### Prediction Interval for a Single Response Value

```{r}
#| echo: true
predict(mod, newdata = data.frame(Illiteracy = 1.4), interval = "prediction", level = 0.95)
```

#### Confidence Interval for a Mean Response Value

```{r}
#| echo: true
predict(mod, newdata = data.frame(Illiteracy = 1.4), interval = "confidence", level = 0.95)
```

[Compare the two intervals. Which one is wider and why do you think this?]{.fragment}

 

[*We won't cover anymore of the details of these intervals, but if you want to read more or explore the formulas used to calculate the interval bounds, check out [the textbook's supplemental material](https://www.openintro.org/go/?id=stat_extra_linear_regression_supp&referrer=os4_pdf).*]{.fragment}

```{r}
library(countdown)
countdown(minutes = 0, seconds = 45)
```

## R Code for This Week's Examples {.smaller}

```{r}
#| echo: true
#| eval: false

# Open the tidyverse library
library(tidyverse)

# Import the dataset, first need to download the data from Canvas
state_30 <- read_csv(file.choose())

# Create a scatterplot of the Illiteracy and LifeExp variables
ggplot(data = state_30, aes(x = Illiteracy, y = LifeExp)) + 
  geom_point(color = "purple", size = 3) + 
  labs(y = "Life Expectancy (years)", 
       x = "Illiteracy Rate (% of population)", 
       title = "Illiteracy Rate vs. Life Expectancy",
       subtitle = "for 30 US States in 1970") + 
  theme(axis.title = element_text(size = 18)) + 
  theme_bw() + 
  stat_smooth(method = "lm",
              formula = y ~ x, 
              geom = "smooth", 
              se = FALSE, 
              color = "darkgreen")

## Calculate the correlation between illiteracy rate and life exp
state_30 %>% summarise(cor = cor(Illiteracy, LifeExp))

# Estimate intercept and slope for LSRL
lm(LifeExp ~ Illiteracy, data = state_30)
```
