---
title: "Topic 10 - Inference for Linear Regression"
format: 
  revealjs:
    theme: OSU.scss
    chalkboard: false
    html-math-method: mathjax
    multiplex: true
editor: visual
---

```{r}
library(tidyverse)
```

```{r}
library(datasets)
 
state <- state.x77 |>
    as_tibble(rownames = "States")

set.seed(7493410)

keep <- sample(1:50, 30, replace = FALSE)

state_30 <- state[keep,]
state_30 <- state_30 |>
  rename("LifeExp"=`Life Exp`)
mod <- lm(LifeExp ~ Illiteracy, data = state_30)

```

## Download or print notes to PDF {.smaller}

If you'd like to export this presentation to a PDF, do the following

::: nonincremental
1.  Toggle into Print View using the E key.
2.  Open the in-browser print dialog (CTRL/CMD+P)
3.  Change the **Destination** to **Save as PDF**.
4.  Change the **Layout** to **Landscape**.
5.  Change the **Margins** to **None**.
6.  Enable the **Background graphics** option.
7.  Click **Save**.
:::

This feature has been confirmed to work in Google Chrome and Firefox.

## Inference for Linear Regression {.smaller}

[Recall that if we're using a **sample** of data to try and model the true relationship between two quantitative variables, then the intercept, $b_0$, and slope, $b_1$, estimates are *random variables*.]{.fragment}

::: columns
::: column
::: fragment
```{r}
#| fig-width: 5
#| fig-height: 4
state_30 |>
  ggplot(aes(x = `HS Grad`, y = Income)) +
  geom_point(color = viridis::viridis(6)[5], size = 3) + 
  labs(y = "Per Capita Income", 
       x = "Percentage of Population with High School Diploma", 
       title = "High School Graduation Rates vs. Income ",
       subtitle = "for 30 US States in 1970") + 
  theme(axis.title = element_text(size = 18)) + 
  theme_bw() + 
  stat_smooth(method = "lm",
              formula = y ~ x, 
              geom = "smooth", 
              se = FALSE, 
              color = viridis::viridis(6)[5])
```
:::

```{r}
mod1 <- lm(Income ~ `HS Grad`, data = state_30)
```

[$\hat{y} = 1994.39 + 46.68 x$]{.fragment}
:::

::: column
```{r}
set.seed(7493489)
keep2 <- sample(1:50, 30, replace = FALSE)

state_30_2 <- state[keep2,]
```

::: fragment
```{r}
#| fig-width: 5
#| fig-height: 4
state_30_2 |>
  ggplot(aes(x = `HS Grad`, y = Income)) +
  geom_point(color = viridis::viridis(6)[2], size = 3) + 
  labs(y = "Per Capita Income", 
       x = "Percentage of Population with High School Diploma", 
       title = "High School Graduation Rates vs. Income ",
       subtitle = "for a different set of 30 US States in 1970") + 
  theme(axis.title = element_text(size = 18)) + 
  theme_bw() + 
  stat_smooth(method = "lm",
              formula = y ~ x, 
              geom = "smooth", 
              se = FALSE, 
              color = viridis::viridis(6)[2])
```
:::

```{r}
mod2 <- lm(Income ~ `HS Grad`, data = state_30_2)
```

[$\hat{y} = 1191.82 + 62.98 x$]{.fragment}
:::
:::

## Estimating Parameters {.smaller}

The LSRL is based on sampled data, so $$\hat{y} = b_0 + b_1x$$ is the estimate for the true population regression equation [$$y = \beta_0 + \beta_1 x + \varepsilon$$]{.fragment}

-   $b_0$ is the point estimate for [$\beta_0$]{.fragment}

-   $b_1$ is the point estimate for [$\beta_1$]{.fragment}

-   $\varepsilon$ is the error (variability around the regression line)

::: {.fragment style="border: 2px solid #2A788EFF; text-align: center"}
Our focus from here on out will be on inference about the slope parameter, $\beta_1$.
:::

## Inference about $\beta_1$ {.smaller}

Consider the LSRL for the relationship between the percentage a state's population with a high school diploma and per capita income.

::: columns
::: column
::: fragment
$\hat{y} = 1994.39 + 46.68 x$

```{r}
#| fig-width: 5
#| fig-height: 4
state_30 |>
  ggplot(aes(x = `HS Grad`, y = Income)) +
  geom_point(color = viridis::viridis(6)[5], size = 3) + 
  labs(y = "Per Capita Income", 
       x = "Percentage of Population with High School Diploma", 
       title = "High School Graduation Rates vs. Income ",
       subtitle = "for 30 US States in 1970") + 
  theme(axis.title = element_text(size = 18)) + 
  theme_bw() + 
  stat_smooth(method = "lm",
              formula = y ~ x, 
              geom = "smooth", 
              se = FALSE, 
              color = viridis::viridis(6)[5])
```
:::
:::

::: column
[The slope in the LSRL, $46.68$, is an **estimate** for the true population regression line.]{.fragment}

[We might wonder, do these data provide strong evidence that the percentage of HS graduates is useful predictor of a state's per capita income?]{.fragment}

::: fragment
Frame this question of interest into a hypothesis test:

$H_0:$ [$\beta_1 = 0$]{.fragment} [The true linear model has slope zero.]{.fragment}

$H_A:$ [$\beta_1 \neq 0$]{.fragment} [The true linear model has a slope different than zero. *The explanatory variable is good predictor of the response.*]{.fragment}
:::
:::
:::

## Using Software Output to Perform a Hypothesis Test on $\beta_1$ {.smaller}

::: fragment
```{r}
state_30 <- state_30 |>
  rename("HSGrad"=`HS Grad`)
```

::: columns
::: column
R Code:

```{r}
#| echo: true

LSRL <- lm(Income ~ HSGrad, data = state_30)
summary(LSRL)
```
:::

::: column
```{r}
#| fig-width: 5
#| fig-height: 4
state_30 |>
  ggplot(aes(x = HSGrad, y = Income)) +
  geom_point(color = viridis::viridis(6)[5], size = 3) + 
  labs(y = "Per Capita Income", 
       x = "Percentage of Population with High School Diploma", 
       title = "High School Graduation Rates vs. Income ",
       subtitle = "for 30 US States in 1970") + 
  theme(axis.title = element_text(size = 18)) + 
  theme_bw() + 
  stat_smooth(method = "lm",
              formula = y ~ x, 
              geom = "smooth", 
              se = FALSE, 
              color = viridis::viridis(6)[5])
```
:::
:::
:::

## Interpreting the p-value and drawing a conclusion {.smaller}

[Perform a hypothesis test on the slope parameter, $\beta_1$, using a significance level of $\alpha = 0.05$.]{.fragment}

[The p-value for the hypothesis test on the slope parameter is $0.000658$.]{.fragment}

[Since $0.000658 < \alpha$, we will [reject]{style="color:red;"} the null hypothesis.]{.fragment}

[There is [convincing]{style="color:red;"} evidence that the percentage of high school graduates is a useful predictor of a state's per capita income in the 1970s.]{.fragment}

## R Code for This Week's Examples {.smaller}

```{r}
#| echo: true
#| eval: false

# Open the tidyverse library
library(tidyverse)

# Import the dataset, first need to download the data from Canvas
state_30 <- read_csv(file.choose())

# Create a scatterplot of the Illiteracy and LifeExp variables
ggplot(data = state_30, aes(x = Illiteracy, y = LifeExp)) + 
  geom_point(color = "purple", size = 3) + 
  labs(y = "Life Expectancy (years)", 
       x = "Illiteracy Rate (% of population)", 
       title = "Illiteracy Rate vs. Life Expectancy",
       subtitle = "for 30 US States in 1970") + 
  theme(axis.title = element_text(size = 18)) + 
  theme_bw() + 
  stat_smooth(method = "lm",
              formula = y ~ x, 
              geom = "smooth", 
              se = FALSE, 
              color = "darkgreen")

## Calculate the correlation between illiteracy rate and life exp
state_30 %>% summarise(cor = cor(Illiteracy, LifeExp))

# Estimate intercept and slope for LSRL
lm(LifeExp ~ Illiteracy, data = state_30)
```


## Practice 🎩 

Please complete the Class 18 Activity in Top Hat.